A robust and memory-efficient transition state search method for complex energy landscapes

### 2022-08-30 manuscript.tex
Jack and Halim's changes to CoS section

% Temporary
\usepackage{soul}
\soulregister\cite7
\soulregister\cref7
\soulregister\num7
\soulregister\multicolumn7
% \renewcommand{\hl}[1]{#1} % Uncomment to remove mark-up

@@

\subsection{Comparison with chain-of-states methods}
\hl{%
In contrast to BITSS, chain-of-states methods do not typically find transition states within a reasonable accuracy.
Instead, they are used to find the full pathway (or an approximation thereof), and a secondary method can then be used to refine to the transition state.
Because these methods can identify multiple transition state candidates, they are advantageous for pathways with many intermediate minima.
However, for systems with few or just a single transition state along the pathways, BITSS offers the possibility for improved efficiency and reduced memory requirements, since just two states are being optimised.
This is particularly likely to be true for complex, non-linear pathways which can need a large number of states to accurately approximate the pathway.
Also, a complicated initial interpolation may be necessary for these pathways, whereas BITSS requires no such initialisation.

Here we compare the speeds of convergence of BITSS to two often-used chain-of-states methods for finding transition states: climbing-image nudged elastic band (CINEB) \cite{Henkelman2000a}, and DNEB with hybrid eigenvector following (DNEB-HEVF) \cite{Cerjan1981}.
It is also possible to use hybrid methods based upon the string method \cite{E2007,Zimmerman2013}, although the results are expected to be similar.
Also, primarily, we aim to observe how the BITSS method behaves for different systems, rather than providing a comprehensive comparison of the available methods, which has been performed in other works \cite{Koslover2007,Sheppard2008}.

The core of these methods involves minimising the total energy of a chain of states, connected by elastic springs to keep them equally spaced along the transition pathway.
CINEB modifies this by altering the behaviour of the state with the highest energy.
The direction of minimisation on this state is inverted in the along pathway direction, effectively converting the saddle point into a local minimum.
Alternatively, DNEB-HEVF involves first minimising the chain of states until a criterion is met, for which we use the root-mean-square of the total energy gradient of the chain. 
At this point eigenvector-following is performed from the highest energy state, moving uphill along the smallest eigenvector of the Hessian until it reaches the transition state.

It is also possible to combine eigenvector-following with BITSS, which we also include in the comparison for comleteness.
For this, the switch from BITSS to eigenvector-following is chosen to occur when the average of the two states changes by less than a tenth of the reduction in the distance given by \cref{eq:diteration}.
}

@@

\hl{%
The results for the three systems are shown in \cref{tab:speed}.
The first thing to note is that using hybrid eigenvector-following does not significantly improve the speed of BITSS.
Indeed, for the wetting example it is actually considerably faster to use BITSS alone.
Next, we see that the chain-of-states methods can be faster than BITSS for the systems with more linear pathways, particularly when using few images.
Although, if too few images are used both CINEB and DNEB-HEVF can struggle, as seen for the cylindrical buckling with three images.
Meanwhile, for the wetting example, CINEB does not converge to the transition state because the estimated tangent vector is highly inaccurate due to the non-linearity of the pathway.
A large number of images would have to be used for it to converge.
Furthermore, we find that BITSS is quite a bit faster than DNEB-HEVF in this case, in line with the prediction that BITSS is advantageous for complex, non-linear pathways.

Overall, while BITSS does not realise the significant improvements in speed that were hypothesised, it is at least comparable with other popular methods, and appears faster for highly non-linear pathways.
It also benefits from lower computational memory demand and not requiring an initial interpolation.
}

@@

These four sections of the phase-field model are obtained using the following equations,
\begin{align}
  E^\mathrm{B}[\phi] &= \sum_i \frac{1}{\epsilon} \left( \frac{{\phi_i}^4}{4} - \frac{{\phi_i}^2}{2} + \frac{1}{4} \right) \upDelta V,\\
  E^\mathrm{I}[\phi] &= \sum_i \frac{\epsilon}{2} \abs{\grad \phi_i}^2 \upDelta V,\\
  E^\mathrm{S}[\phi] &= \sum_j \sqrt{2}\cos\theta_j \left( \frac{{\phi_j}^3}{6} - \frac{\phi_j}{2} - \frac{1}{3} \right) \upDelta S,\\
  E^\mathrm{V}[\phi] &= k_\mathrm{V} \left[ \sum_i \frac{\phi_i + 1}{2} \upDelta V - V_0 \right]^2,
\end{align}
where the index $i$ includes all of the nodes, while $j$ represents the nodes along the solid surface.
$\upDelta V$ and $\upDelta S$ are, respectively, the volume and solid surface areas contained by each individual node.
$\epsilon$ is the liquid-gas interface width (set to 2.5 lattice units), $\theta_j$ is the contact-angle with the solid surface, and $V_0$ is the constrained volume of the liquid drop.
The strength of the volume constraint is parametrised by $k_\mathrm{V}$ for which we use a value of $10^4$.
The gradients of these terms are given by,
\begin{align}
  \frac{\partial E^\mathrm{B}}{\partial \phi_i} &= \frac{1}{\epsilon} \left( {\phi_i}^3 - \phi_i \right) \upDelta V,\\
  \frac{\partial E^\mathrm{I}}{\partial \phi_i} &= \frac{\epsilon}{2} \left[
    \frac{\partial \abs{\grad \phi_i}^2}{\partial \phi_i} +
    \sum_k\frac{\partial \abs{\grad \phi_k}^2}{\partial \phi_i} \right] \upDelta V,\\
  \frac{\partial E^\mathrm{S}}{\partial \phi_j} &= \sqrt{2}\cos\theta_j \left( \frac{{\phi_j}^2}{2} - \frac{1}{2} \right) \upDelta S,\\
  \frac{\partial E^\mathrm{V}}{\partial \phi_i} &= k_\mathrm{V} \left[ \sum_{i'} \frac{\phi_{i'} + 1}{2} \upDelta V - V_0 \right] \upDelta V,
\end{align}
where index $k$ denotes neighboring nodes at which the evaluation of the gradient uses $\phi_i$.

>>>>

% Temporary
\usepackage{xcolor}
\usepackage{soul}
\soulregister\cite7
\soulregister\cref7
\soulregister\num7
\soulregister\multicolumn7
% \renewcommand{\hl}[1]{#1} % Uncomment to remove mark-up

@@

\subsection{Comparison with chain-of-states methods}
\hl{%
In contrast to BITSS, chain-of-states methods do not typically find transition states directly.
Instead, they are designed to find the full pathway (or an approximation thereof), and a secondary method can then be used to refine to the transition state.
As we will demonstrate, this strategy is successful for simple, linear pathways, but faces two key challenges when the pathway is highly non-linear.
Firstly, for such complex pathways, a large number of states are required to sufficiently approximate the minimum energy pathway.
The second is that choosing a suitable initial interpolation can be problematic to achieve.
BITSS is advantageous in both these regards, as only two states are evolved, regardless of the pathway complexity, and no initial interpolation is required.
For these advantages to be of practical use, BITSS must demonstrate at least a comparable speed of convergence to the transition states, relative to these hybrid chain-of-states methods, which we now investigate.

Here we compare the speeds of convergence of BITSS to two widely used approaches for finding transition states that employ chain-of-states methods: climbing-image nudged elastic band (CINEB) \cite{Henkelman2000a}, and DNEB with hybrid eigenvector following (DNEB-HEVF) \cite{Cerjan1981}.
}{\color{blue}
It is also possible to use the string method with a climbing image \cite{E2007} or eigenvector following \cite{Zimmerman2013}; although, in this case, the results are expected to be similar to the nudged-elastic-band methods.
}\hl{%
The core of these methods involves minimising the total energy of a chain of states, connected by elastic springs to keep them equally spaced along the transition pathway.
CINEB modifies this by altering the behaviour of the state with the highest energy.
The direction of minimisation on this state is inverted along the pathway direction, effectively converting the saddle point into a local minimum.
Alternatively, DNEB-HEVF involves first minimising the chain of states until a criterion is met, for which we use the root-mean-square of the total energy gradient of the chain.
At this point eigenvector-following is performed from the highest energy state, moving uphill along the smallest eigenvector of the Hessian until it reaches the transition state.
It is also possible to combine eigenvector-following with BITSS, which we also include in the comparison for completeness.
}{\color{red}
Move this to methods: For this, the switch from BITSS to eigenvector-following is chosen to occur when the average of the two states changes by less than a tenth of the reduction in the distance given by \cref{eq:diteration}.
}\hl{%
We note that our aim in this section is to observe how the BITSS method behaves for different systems, rather than providing a comprehensive comparison of the currently available methods, which has been performed in other works \cite{Koslover2007,Sheppard2008}.
}

@@

\hl{%
The results for the three systems are shown in \cref{tab:speed}.
First, we note that for all three systems, using hybrid eigenvector-following does not significantly improve the speed of BITSS.
Indeed, for buckling and especially wetting, HEVF is detrimental to performance.
Next, it is interesting to compare each method's performance between simple and complex pathways.
In contrast to the wetting transition's highly non-linear pathway, the pathways of the LJ-7 rearrangement and the buckling system can be simply tracked following a gradual variation in the order parameters.
For LJ-7, this is the translation of atoms 5 and 6, and for buckling, this is the radial displacement of the centre of the dimple \cite{Panter2019}.
For the simpler pathways, BITSS is generally slower, but for the complex pathways, BITSS is faster.
Moreover, we see that for the wetting example, CINEB does not converge to the transition state because the estimated tangent vector is highly inaccurate due to the non-linearity of the pathway.

For situations where memory is limited, it is important to minimise the number of images used.
As is shown for cylindrical buckling with three images, efficiently finding the TS is challenging for both CINEB and DNEB-HEVF.
BITSS however converges using only two images.

However, if too few images are used (the minimum number of which is system dependant), both CINEB and DNEB-HEVF can struggle, as observed for the cylindrical buckling with three images.
}

@@

These four sections of the phase-field model are obtained using the following equations,
\begin{align}
  E^\mathrm{B}[\phi] &= \sum_i \frac{1}{\epsilon} \left( \frac{{\phi_i}^4}{4} - \frac{{\phi_i}^2}{2} + \frac{1}{4} \right) \upDelta V,\\
  E^\mathrm{I}[\phi] &= \sum_i \frac{\epsilon}{2} \abs{\grad \phi_i}^2 \upDelta V,\\
  E^\mathrm{S}[\phi] &= \sum_j \sqrt{2}\cos\theta_j \left( \frac{{\phi_j}^3}{6} - \frac{\phi_j}{2} - \frac{1}{3} \right) \upDelta S,\\
  E^\mathrm{V}[\phi] &= k_\mathrm{V} \left[ \sum_i \frac{\phi_i + 1}{2} \upDelta V - V_0 \right]^2,
\end{align}
where the index $i$ includes all of the nodes, while $j$ represents the nodes along the solid surface.
$\upDelta V$ and $\upDelta S$ are, respectively, the volume and solid surface areas contained by each individual node.
$\epsilon$ is the liquid-gas interface width (set to 2.5 lattice units), $\theta_j$ is the contact-angle with the solid surface, and $V_0$ is the constrained volume of the liquid drop.
The strength of the volume constraint is parametrised by $k_\mathrm{V}$ for which we use a value of $10^4$.
The gradients of these terms are given by,
\begin{align}
  \frac{\partial E^\mathrm{B}}{\partial \phi_i} &= \frac{1}{\epsilon} \left( {\phi_i}^3 - \phi_i \right) \upDelta V,\\
  \frac{\partial E^\mathrm{I}}{\partial \phi_i} &= \frac{\epsilon}{2} \left[
    \frac{\partial \abs{\grad \phi_i}^2}{\partial \phi_i} +
    \sum_k\frac{\partial \abs{\grad \phi_k}^2}{\partial \phi_i} \right] \upDelta V,\\
  \frac{\partial E^\mathrm{S}}{\partial \phi_j} &= \sqrt{2}\cos\theta_j \left( \frac{{\phi_j}^2}{2} - \frac{1}{2} \right) \upDelta S,\\
  \frac{\partial E^\mathrm{V}}{\partial \phi_i} &= k_\mathrm{V} \left[ \sum_{i'} \frac{\phi_{i'} + 1}{2} \upDelta V - V_0 \right] \upDelta V,
\end{align}
where index $k$ denotes neighbouring nodes at which the evaluation of the gradient uses $\phi_i$.

{\color{red}
\section{Implementation of the hybrid methods, CINEB, DENB-HEVF, and BITSS-HEVF}
JRP - Sam can you fill out this section to briefly detail the method parameters?
Also, say that HEVF was performed using the program OPTIM, with the reference http://www-wales.ch.cam.ac.uk/software.html.
}
